---
title: "Homework-2"
author: "waheeb Algabri"
output:
  html_document:
    highlight: pygments
    theme: cerulean
    toc: true
    toc_float: true
  pdf_document: default
editor_options: 
  chunk_output_type: console
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(eval = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(openintro)
```



# Problem Set Solutions in Linear Algebra

### Problem Set 1: Matrix Non-commutativity and Symmetric Matrices

Objective: Show that $ATA ≠ $AAT$in general and discuss conditions under which $ATA = AAT$

Matrix multiplication is non-commutative, meaning that the order of multiplication matters. For instance, given a matrix *A*:


A * \cdot A^T = \begin{bmatrix}
    a_{11} & a_{12} & \dots & a_{1n} \\
    a_{21} & a_{22} & \dots & a_{2n} \\
    \vdots & \vdots & \ddots & \vdots \\
    a_{m1} & a_{m2} & \dots & a_{mn}
\end{bmatrix}
\cdot
\begin{bmatrix}
    a_{11} & a_{21} & \dots & a_{m1} \\
    a_{12} & a_{22} & \dots & a_{m2} \\
    \vdots & \vdots & \ddots & \vdots \\
    a_{1n} & a_{2n} & \dots & a_{mn}
\end{bmatrix}

***
$(A^T A)_{ij} = \sum_{k=1}^{n} a_{ik} a_{jk}$


It is evident that if $n \neq m$ then $A^TA$and $AA^T$ cannot be equal due to their different dimensions.

Symmetric matrices include diagonal square matrices, such as the identity matrix **I**. For instance:

$$I = \begin{bmatrix}
  1 & 0 & 0 \\
  0 & 1 & 0 \\
  0 & 0 & 1 \\
\end{bmatrix}$$




For example, suppose we have $A$ and $A^T$ as follows:

Let $B$ be a matrix given by:
\[ B = \begin{bmatrix} 1 & 2 & 3 \\ 4 & 5 & 6 \end{bmatrix} \]

The transpose of $B$, denoted as $B^T$, is:
\[ B^T = \begin{bmatrix} 1 & 4 \\ 2 & 5 \\ 3 & 6 \end{bmatrix} \]

Now, let's find the product $BB^T$:
\[ BB^T = \begin{bmatrix} 1 & 2 & 3 \\ 4 & 5 & 6 \end{bmatrix} \begin{bmatrix} 1 & 4 \\ 2 & 5 \\ 3 & 6 \end{bmatrix} \]

The element at position $(i, j)$ in $BB^T$ is given by:
\[ (BB^T)_{ij} = \sum_{k=1}^{3} b_{ik} \cdot (b_{jk})^T \]

After performing the calculations, the result is a $2 \times 2$ matrix:
\[ BB^T = \begin{bmatrix} 14 & 32 \\ 32 & 77 \end{bmatrix} \]

Next, let's find the product $B^TB$:
\[ B^TB = \begin{bmatrix} 1 & 4 \\ 2 & 5 \\ 3 & 6 \end{bmatrix} \begin{bmatrix} 1 & 2 & 3 \\ 4 & 5 & 6 \end{bmatrix} \]

The element at position $(i, j)$ in $B^TB$ is given by:
\[ (B^TB)_{ij} = \sum_{k=1}^{3} (b_{ik})^T \cdot b_{kj} \]

After performing the calculations, the result is a $3 \times 3$ matrix:
\[ B^TB = \begin{bmatrix} 17 & 22 & 27 \\ 22 & 29 & 36 \\ 27 & 36 & 45 \end{bmatrix} \]

As in the previous example, $BB^T$ and $B^TB$ are not equal. Therefore, in general, $BB^T \neq B^TB$ as $B^T \neq B$.
***

*For a special type of square matrix we get ATA=AAT. Under what conditions could this be true? (Hint: The Identity matrix I is an example of such a matrix)*


Let $C$ be a square matrix such that $C^T = C$. This means that $C$ is a symmetric matrix.

\[ C = \begin{bmatrix} c_{11} & c_{12} & \dots & c_{1n} \\ c_{21} & c_{22} & \dots & c_{2n} \\ \vdots & \vdots & \ddots & \vdots \\ c_{n1} & c_{n2} & \dots & c_{nn} \end{bmatrix} \]

Now, let's find $C^TC$ and $CC^T$:

1. $C^TC$:
\[ C^TC = \begin{bmatrix} c_{11} & c_{21} & \dots & c_{n1} \\ c_{12} & c_{22} & \dots & c_{n2} \\ \vdots & \vdots & \ddots & \vdots \\ c_{1n} & c_{2n} & \dots & c_{nn} \end{bmatrix} \begin{bmatrix} c_{11} & c_{12} & \dots & c_{1n} \\ c_{21} & c_{22} & \dots & c_{2n} \\ \vdots & \vdots & \ddots & \vdots \\ c_{n1} & c_{n2} & \dots & c_{nn} \end{bmatrix} \]

The element at position $(i, j)$ in $C^TC$ is given by:
\[ (C^TC)_{ij} = \sum_{k=1}^{n} c_{ki} \cdot c_{kj} \]

2. $CC^T$:
\[ CC^T = \begin{bmatrix} c_{11} & c_{12} & \dots & c_{1n} \\ c_{21} & c_{22} & \dots & c_{2n} \\ \vdots & \vdots & \ddots & \vdots \\ c_{n1} & c_{n2} & \dots & c_{nn} \end{bmatrix} \begin{bmatrix} c_{11} & c_{21} & \dots & c_{n1} \\ c_{12} & c_{22} & \dots & c_{n2} \\ \vdots & \vdots & \ddots & \vdots \\ c_{1n} & c_{2n} & \dots & c_{nn} \end{bmatrix} \]

The element at position $(i, j)$ in $CC^T$ is given by:
\[ (CC^T)_{ij} = \sum_{k=1}^{n} c_{ik} \cdot c_{jk} \]

Since $C^T = C$, the elements of $C^TC$ and $CC^T$ will be the same. Therefore, in this case, $C^TC = CC^T$ for any symmetric square matrix $C$. This includes diagonal matrices and, in particular, the identity matrix $I$.


### Problem Set 2: LU Factorization Function in R

*Define the LU or LDU Factorization Function*

```{r}
luOrLduFactorize <- function(A, ldu = FALSE) {
  n <- nrow(A)
  U <- A
  L <- diag(n)
  D <- diag(n)
  
  for (i in 1:(n - 1)) {
    for (j in (i + 1):n) {
      L[j, i] = U[j, i] / U[i, i]
      U[j, ] = U[j, ] - L[j, i] * U[i, ]
      
      if (ldu) {
        D[j, j] = U[j, j]
        U[j, j] = 1  # Set diagonal elements of U to 1 in LDU factorization
      }
    }
  }
  
  if (ldu) {
    result <- list("L" = L, "D" = D, "U" = U)
  } else {
    result <- list("L" = L, "U" = U)
  }
  
  return(result)
}

```



*Factorize a Square Matrix A into LU*
```{r}
A <- matrix(c(1, 2, 3, 1, 1, 1, 2, 0, 1), nrow = 3)

luResult <- luOrLduFactorize(A)

# Check if LU factorization is correct
if (all(luResult$L %*% luResult$U == A)) {
  cat("LU Factorization is correct:\n")
  print("Original Matrix A:")
  print(A)
  print("Lower Triangular Matrix L:")
  print(luResult$L)
  print("Upper Triangular Matrix U:")
  print(luResult$U)
} else {
  cat("LU Factorization is incorrect.\n")
}

# Factorize a Square Matrix A into LDU
A <- matrix(c(1, 2, 3, 1, 1, 1, 2, 0, 1), nrow = 3)

lduResult <- luOrLduFactorize(A, ldu = TRUE)

```


*Check if LDU factorization is correct*
```{r}
if (all(lduResult$L %*% lduResult$D %*% lduResult$U == A)) {
  cat("LDU Factorization is correct:\n")
  print("Original Matrix A:")
  print(A)
  print("Lower Triangular Matrix L:")
  print(lduResult$L)
  print("Diagonal Matrix D:")
  print(lduResult$D)
  print("Upper Triangular Matrix U:")
  print(lduResult$U)
} else {
  cat("LDU Factorization is incorrect.\n")
}
```



