---
title: "Homework-3"
author: "waheeb Algabri"
output:
  html_document:
    highlight: pygments
    theme: cerulean
    toc: true
    toc_float: true
  pdf_document: default
editor_options: 
  chunk_output_type: console
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(eval = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(openintro)
library(pracma)
library(Matrix)
```


### Problem set 1

**1-What is the rank of the matrix A?**

\[
A = \begin{bmatrix}
1 & 2 & 3 & 4 \\
-1 & 0 & 1 & 3 \\
0 & 1 & -2 & 1 \\
5 & 4 & -2 & -3
\end{bmatrix}
\]

As we know that the Nonsingular Matrix Equivalence Theorem, often denoted as NME8, states that for a square matrix $A$

the following statements are equivalent:

1- $A$ is nonsingular (invertible).

2- rank $(A) = n$ , where n is the number of rows (or columns) of $A$

3- The determinant of A is nonzero $(det(A ≠ 0)$

4- The columns (or rows) of are linearly independent

5- The reduced row echelon form of $A$ is the identity matrix.


```{r}
# Define matrix A
A <- matrix(c(1, 2, 3, 4, 
              -1, 0, 1, 3, 
              0, 1, -2, 1, 
              5, 4, -2, -3), 
            nrow = 4, byrow = TRUE)

# Compute the determinant of A
det_A <- det(A)

# Compute the rank of A
rank_A <- Matrix::rankMatrix(A)

# Check if A is nonsingular (invertible) using NME8
if (det_A != 0) {
  cat("Matrix A is nonsingular.\n")
  cat("Rank of A:", rank_A, "\n")
} else {
  cat("Matrix A is singular.\n")
}

```

***

**2- Given an mxn matrix where $m > n$, what can be the maximum rank? The mini- mum rank, assuming that the matrix is non-zero?**

For an $m × n$ matrix where $m > n$ , the maximum rank that can be achieved is $n$, which is the smaller dimension. This is because the rank of a matrix is bounded by the number of columns or rows, whichever is smaller. Since in this case,$n$ is smaller, the maximum rank cannot exceed n.

On the other hand, the minimum rank of the matrix, assuming it is non-zero, is 1. This is because a non-zero matrix will always have at least one linearly independent row or column, which implies that the rank cannot be zero, So,Maximum rank: $n$ and Minimum rank: $1$

```{r}
# Example of an m x n matrix where m > n
m <- 4  # number of rows
n <- 2  # number of columns

# Create a random non-zero matrix of size m x n
set.seed(123)  # for reproducibility
example_matrix <- matrix(runif(m * n, min = -5, max = 5), nrow = m)

# Display the matrix
print(example_matrix)

# Compute the rank of the matrix
rank_example <- Matrix::rankMatrix(example_matrix)
print(paste("Rank of the matrix:", rank_example))

```

The maximum rank that can be achieved is 2 because $n = 2$, which is the smaller dimension of the matrix. as computed, the rank of the matrix is indeed 2

This confirms the theoretical expectation that the maximum rank for this matrix, where $m > n$

***
**3- What is the rank of matrix B?**

\begin{bmatrix}
1 & 2 & 1 \\
3 & 6 & 3 \\
2 & 4 & 2
\end{bmatrix}


```{r}
# Define matrix B
B <- matrix(c(1, 2, 1,
              3, 6, 3,
              2, 4, 2), 
            nrow = 3, byrow = TRUE)

# Compute the rank of B
rank_B <- Matrix::rankMatrix(B)
print(paste("Rank of matrix B:", rank_B))

```



### Problem set 2

Compute the eigenvalues and eigenvectors of the matrix A. You’ll need to show your
work. You’ll need to write out the characteristic polynomial and show your solution.


$A = \begin{bmatrix}
1 & 2 & 3 \\
0 & 4 & 5 \\
0 & 0 & 6
\end{bmatrix}$



*Solution :*

```{r}
A <- matrix(c(1, 2, 3, 0, 4, 5, 0, 0, 6), nrow = 3, byrow = TRUE)
print(A)
```

Since A is a *triangular* matrix, its eigenvalues are values on the diagonal, so 

$λ1 = 1$, $λ2 = 4$, and λ3 = 6

Double-check eigenvalues in R
```{r}
eigen(A)$values
```

The characteristic polynomial is $p_A(\lambda) = (1 - \lambda)(4 - \lambda)(6 - \lambda)$ or 

$p_A(\lambda) = 24 - 34\lambda + 11\lambda^2 - \lambda^3$.

\subsection*{Eigenvalue $\lambda = 1$:}

If $\lambda = 1$, then $A - \lambda I_3$ is row-reduced to:
```{r}
rref(A - 1 * diag(3))
```


\[
\begin{pmatrix}
0 & 1 & 0 \\
0 & 0 & 1 \\
0 & 0 & 0 \\
\end{pmatrix}
\]

This gives us the *eigenspace8 $E_{\lambda=1} = \langle \begin{pmatrix} 1 \\ 0 \\ 0 \end{pmatrix} \rangle$.

\subsection*{Eigenvalue $\lambda = 4$:}

If $\lambda = 4$, then $A - \lambda I_3$ is row-reduced to:
```{r}
rref(A - 4 * diag(3))
```


\[
\begin{pmatrix}
1 & -0.6666667 & 0 \\
0 & 0 & 1 \\
0 & 0 & 0 \\
\end{pmatrix}
\]

This gives us the *eigenspace* $E_{\lambda=4} = \langle \begin{pmatrix} 1 \\ 1.5 \\ 0 \end{pmatrix} \rangle$.

\subsection*{Eigenvalue $\lambda = 6$:}

If $\lambda = 6$, then $A - \lambda I_3$ is row-reduced to:
```{r}
rref(A - 6 * diag(3))
```


\[
\begin{pmatrix}
1 & 0 & -1.6 \\
0 & 1 & -2.5 \\
0 & 0 & 0 \\
\end{pmatrix}
\]

This gives us the *eigenspace* $E_{\lambda=6} = \langle \begin{pmatrix} 1.6 \\ 2.5 \\ 1 \end{pmatrix} \rangle$.


