---
title: "Homework-3"
author: "waheeb Algabri"
output:
  html_document:
    highlight: pygments
    theme: cerulean
    toc: true
    toc_float: true
  pdf_document: default
editor_options: 
  chunk_output_type: console
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(eval = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(openintro)
library(pracma)
library(Matrix)
```


## Loading Necessary Libraries:
```{r}
# Setup: Loading necessary libraries
library(jpeg)
library(EBImage)
library(OpenImageR)
library(tidyverse)
library(foreach)
```


**With the attached data file, build and visualize eigenimagery that accounts for 80% of the variability. Provide full R code and discussion.**

## Loading in the Images

```{r}

# Set the path to the directory containing JPEG files
path <- "~/Desktop/Data-605/Data-605/Homeworks/jpg"

# List all JPEG files in the specified directory
files <- list.files(path, pattern = "\\.jpg", full.names = TRUE)

# Calculate the number of JPEG files
num <- length(files)

# Define image dimensions
height <- 1200
width <- 2500

# Set the scaling factor
scale <- 20

# Calculate new dimensions after scaling
new_height <- height / scale
new_width <- width / scale

```


## Load the Data into an Array and Vectorize

Since the images are very large, we resize each image according to a chosen scale and then load each image into an array of  scaled dimension. 


```{r}
# Create an array to store resized images
array_a <- array(rep(0, num * new_height * new_width * 3), dim = c(num, new_height, new_width, 3))

# Resize and store each image in the array
for (i in 1:num) {
    temp <- resizeImage(readJPEG(files[i]), new_height, new_width)
    array_a[i, , , ] <- array(temp, dim = c(1, new_height, new_width, 3))
}

```

We then create a matrix of the RGB components of each image by looping through the array of scaled images. Within the loop, the R, G, and B components of each image are converted to vectors. These vectors are then concatenated and transposed and added into the matrix of images. 

```{r}
# Flatten the array and store pixel values in a data frame
flat <- matrix(0, num, prod(dim(array_a)))

for (i in 1:num) {
    r <- as.vector(array_a[i, , , 1])
    g <- as.vector(array_a[i, , , 2])
    b <- as.vector(array_a[i, , , 3])
    flat[i, ] <- t(c(r, g, b))
}

# Convert the flattened matrix to a data frame
shoes <- as.data.frame(t(flat))

```


## Creating a Function to Plot the Images

```{r}
# Define a function to plot JPEG images
plot_jpeg <- function(path, add = FALSE) {
    jpg <- readJPEG(path, native = TRUE)
    res <- dim(jpg)[2:1]
    if (!add) {
        plot(1, 1, xlim = c(1, res[1]), ylim = c(1, res[2]), asp = 1, type = "n",
            xaxs = "i", yaxs = "i", xaxt = "n", yaxt = "n", xlab = "", ylab = "",
            bty = "n")
    }
    rasterImage(jpg, 1, 1, res[1], res[2])
}

# Plot resized images
par(mfrow = c(3, 3))
par(mai = c(0.3, 0.3, 0.3, 0.3))
for (i in 1:num) {
    plot_jpeg(writeJPEG(array_a[i, , , ]))
}

```

## Standardize the Pixel Values
```{r}
# Standardize the pixel values
scaled <- scale(shoes, center = TRUE, scale = TRUE)

# Extract mean and standard deviation of standardized pixel values
mean.shoe <- attr(scaled, "scaled:center")
std.shoe <- attr(scaled, "scaled:scale")

```


## Calculate the Correlation Matrix
```{r}
# Calculate the correlation matrix
Sigma_ <- cor(scaled)

```


## Compute Eigenvalues of the Correlation Matrix
```{r}
# Compute eigenvalues of the correlation matrix
myeigen <- eigen(Sigma_)
eigenvalues <- myeigen$values
eigenvalues

```


## extract Eigenvectors of the Correlation Matrix:
```{r}
# Extract eigenvectors of the correlation matrix
eigenvectors <- myeigen$vectors

# Display the first few rows of eigenvectors in a nice table format
knitr::kable(head(eigenvectors), format = "markdown")

```


## Read and Plot a Sample JPEG Image:
```{r}
# Read and plot a sample JPEG image
testing_img <- readJPEG(files[5])
plot(1:2, type = "n", main = "")
rasterImage(testing_img, 1, 1, 2, 2)

```

## Perform PCA on Standardized Pixel Values


```{r}
# Perform PCA on standardized pixel values
scaling <- diag(eigenvalues[1:5]^(-1/2)) / (sqrt(nrow(scaled) - 1))
eigenshoes <- scaled %*% eigenvectors[, 1:5] %*% scaling

# Visualize the fourth eigenshoe
par(mfrow = c(2, 3))
imageShow(array(eigenshoes[, 5], c(new_height, new_width, 3)))

```



## Calculate Cumulative Variance Explained by Each Eigenshoe: 
```{r}
# Calculate cumulative variance explained by each eigenshoe
cumulative_variance <- cumsum(myeigen$values) / sum(myeigen$values)
cumulative_variance

```


## Determine the Number of Principal Components Explaining at Least 80% of the Variance:
```{r}
# Determine the number of principal components explaining at least 80% of the variance
num_comp <- which(cumulative_variance >= 0.8)[1]
num_comp

```

## Plot Cumulative Variance Explained by Eigenshoes: 
```{r}
# Plot cumulative variance explained by eigenshoes
ggplot(as.data.frame(cumulative_variance), aes(x = 1:num, cumulative_variance)) +
    geom_line() + geom_point() + labs(x = "Number of eigenshoes", y = "Cumulative Variance") +
    scale_x_continuous(breaks = seq(1, 17, by = 2)) + theme_minimal()

```







