---
title: "Discussion 3 "
author: "waheeb Algabri"
output:
  html_document:
    highlight: pygments
    theme: cerulean
    toc: true
    toc_float: true
  pdf_document: default
editor_options: 
  chunk_output_type: console
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(eval = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(openintro)
```


### Exercise C20† Chapter E


**Consider the matrix A below. First, show that A is diagonalizable by computing the geometric multiplicities of the eigenvalues and quoting the relevant theorem. Second, find a diagonal matrix D and a nonsingular matrix S so that S−1AS = D. (See Exercise EE.C20 for some of the necessary computations.)**


Consider the matrix $A$:

\[
A = \begin{bmatrix}
    18 & -15 & 33 & -15 \\
    -4 & 8 & -6 & 6 \\
    -9 & 9 & -16 & 9 \\
    5 & -6 & 9 & -4
\end{bmatrix}
\]

\subsection*{Step 1: Diagonalizability and Eigenvalues}

\subsubsection*{1. Find Eigenvalues:}

### Solution 
To solve the Question first let us set up the characteristic polynomial: $\text{det}(A - \lambda I) = 0$.

\[
\text{det}\left(\begin{bmatrix}
    18-\lambda & -15 & 33 & -15 \\
    -4 & 8-\lambda & -6 & 6 \\
    -9 & 9 & -16-\lambda & 9 \\
    5 & -6 & 9 & -4-\lambda
\end{bmatrix}\right) = 0
\]

 To solve for eigenvalues $\lambda$. The characteristic polynomial is given by:

\[
\lambda^4 - 6\lambda^3 - 59\lambda^2 + 250\lambda - 300
\]

Now, I,m going to find the roots of this polynomial to get the eigenvalues.


```{r}
if (!requireNamespace("pracma", quietly = TRUE)) {
  install.packages("pracma")
}

# Load the pracma package
library(pracma)

# Given matrix A
matrixA <- matrix(c(18, -15, 33, -15, -4, 8, -6, 6, -9, 9, -16, 9, 5, -6, 9, -4), 4, 4)

# Step 1: Find eigenvalues and eigenvectors
eigen_result <- eigen(matrixA)

# Extract eigenvalues and eigenvectors
eigenvalues <- eigen_result$values
eigenvectors <- eigen_result$vectors

# Step 2: Check algebraic and geometric multiplicities
for (i in 1:length(eigenvalues)) {
  lambda <- eigenvalues[i]
  algebraic_multiplicity <- sum(eigenvalues == lambda)
  
  # Calculate eigenspace using nullspace from pracma package
  eigenspace <- nullspace(matrixA - lambda * diag(nrow(matrixA)))
  geometric_multiplicity <- ncol(eigenspace)
  
  print(paste("Eigenvalue:", lambda))
  print(paste("Algebraic Multiplicity:", algebraic_multiplicity))
  print(paste("Geometric Multiplicity:", geometric_multiplicity))
  
  # Form matrix S using eigenvectors
  if (geometric_multiplicity > 1) {
    eigenspace_matrix <- eigenvectors[, seq_len(geometric_multiplicity)]
    print(eigenspace_matrix)
  }
}

# Step 3: Check diagonalizability and form matrix S
S <- eigenvectors
D <- diag(eigenvalues)

# Step 4: Verify diagonalization
eigen_sol <- round(solve(S) %*% matrixA %*% S, 2)
print(eigen_sol)

```



### Conclusion 

The matrix A has been shown to be diagonalizable through the computation of its eigenvalues and corresponding eigenvectors. The algebraic and geometric multiplicities of each eigenvalue were carefully examined, confirming that A is diagonalizable. The diagonal matrix D and nonsingular matrix S satisfying the relationship $S^(-1)AS = D$ were successfully determined. This establishes the diagonalization of matrix A 


