---
title: "Discussion-2"
author: "waheeb Algabri"
output:
  html_document:
    highlight: pygments
    theme: cerulean
    toc: true
    toc_float: true
  pdf_document: default
editor_options: 
  chunk_output_type: console
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(eval = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(openintro)
```

**C30 For the matrix A below, compute the dimension of the null space of A, dim (N(A)).**


```{r code_block, echo=FALSE}
A = matrix(c(1,3,0,1),2,2)
B = matrix(c(5,3,1,4),2,2)
```


$$
\begin{bmatrix} 
2 & -1 & -3 & 11 & 9 \\ 
1 & 2 & 1 & -7 & -3 \\ 
3 & 1 & -3 & 6 & 8 \\
2 & 1 & 2 & -5 & -3 \\
\end{bmatrix}

$$





```{r}
A <- matrix(c(2, -1, -3, 11, 9,
              1, 2, 1, -7, -3,
              3, 1, -3, 6, 8,
              2, 1, 2, -5, -3), 
            nrow = 4, byrow = TRUE)

```

```{r}
print(A)
```
To find the dimension of the null space of matrix $$A (dim(N(A)))$$ we need to perform row reduction **(Gaussian elimination)** on the augmented matrix $$[A | 0]$$The null space of A is the set of all solutions to the homogeneous system $$Ax = 0$$.

```{r}
# row reduce A
library(pracma)
print(rref(A))
```
This matrix indicates that there are three pivot columns **(columns 1, 2, and 3)** and two free columns **(columns 4 and 5)** The rank of **A** is the number of pivot columns, which is 3 So, the dimension of the null space $$(dim(N(A))$$ is given by the difference between the total number of columns in **A** and the rank of **A** Therefore: Dimension of the null space $$(dim(N(A)) = Total columns − Rank of A$$

Dimension of the null space $$(dim(N(A)) = 5 −3 = 2$$
So, the dimension of the null space of matrix = 2

